Switch equivalence and weak prime domains for left-linear DPO rewriting systems

### 2025-01-11 full_version/permutations.tex
fine permutations

\begin{definition}Let $(\X, \R)$ be a left-linear DPO-rewriting system,  an \emph{abstraction equivalence} between two derivations $\der{D}$ and $\der{D'}$ with the same length and such that $r(\der{D})=r(\der{D}')$, is a family of isomorphisms $\{\phi_X\}_{X\in \Deltamin(\der{D})}$ such that, for every $i\in [0, \lgh(\der{D})]$ the following diagram commutes
	\[\xymatrix@C=40pt{G'_i&D'_i \ar[r]^{g'_i} \ar@{>->}[l]_{f'_i}&G'_{i+1}\\  L_i \ar[u]^{n'_i} \ar[d]_{n_i}& K_i \ar[u]^{k'_i} \ar[d]_{k_i} \ar[r]^{r_i} \ar@{>->}[l]_{l_i} &R\ar[u]^{h'_i} \ar[d]_{h_i}\\G_i \ar@/_.45cm/[uu]_(.35){\phi_{G_i}}|\hole&D_i\ar@{>->}[l]^{f_i}\ar@/_.45cm/[uu]_(.35){\phi_{D_i}}|\hole \ar[r]_{g_i}&G_{i+1}\ar@/_.45cm/[uu]_{\phi_{H_i}}}\]

	Given two decorated derivations $(\der{D}, \alpha, \omega)$ and $(\der{D}', \alpha', \omega')$, we say that they are \emph{abstraction equivalent}, if $\lgh(\der{D})=\lgh(\der{D}')$, $r(\der{D})=r(\der{D'})$, and there exists an of abstraction equivalence between $\der{D}$ and  $\der{D}$ such that the triangles below commute.
	\[\xymatrix@C=15pt{&\pi(G_0) \ar[dr]^{\alpha'} \ar[dl]_{\alpha}&&& \pi(G_{n+1}) \ar[dr]^{\omega'} \ar[dl]_{\omega}\\ G_0 \ar[rr]_{\phi_{G_0}} && G'_0 &G_{n+1} \ar[rr]_{\phi_{G_{n+1}}} && G'_{n+1} } \]
	We will use $\equiv^a$ to denote the resulting relation.
\end{definition}

\begin{remark}\label{rem:equi}
	It is immediate to see that $\equiv^a$ is an equivalence relation. Indeed $(\der{D}, \alpha, \omega)$ is abstract equivalent to itself via the abstract equivalence with the identities as components. Furthermore, if  $\{\phi_X\}_{X\in \Deltamin(\der{D})}$  witnesses $(\der{D}, \alpha, \omega)\equiv^a (\der{D}', \alpha', \omega')$, then considering $\{\phi^{-1}_X\}_{X\in \Deltamin(\der{D})}$ shows $(\der{D}', \alpha', \omega)\equiv^a (\der{D}, \alpha, \omega)$. Finally, transitivity is assured composing abstract equivalences.

	We will denote by $[\der{D}, \alpha, \omega]_a$ is just the equivalence class of  $(\der{D}, \alpha, \omega)$.
	Such equivalence classes will be called  \emph{abstract decorated derivation}.  
\end{remark}

\begin{example}\label{rem:empty}
	Let $(\der{D},\alpha, \omega)$ be an empty derivation from an object $G$ and  $(\der{D}',\alpha', \omega')$ a another empty one from $G'$.  If $(\der{D},\alpha, \omega)\equiv^a(\der{D}',\alpha', \omega')$ then an abstraction equivalence between $\der{D}$ and $\der{D}'$ is just an isomorphism $\phi\colon G\to G'$, so that $\pi(G)=\pi(G')$. Moreover, such isomorphism must fit in the diagrams below.
	\[\xymatrix@C=15pt{&\pi(G) \ar[dr]^{\alpha'} \ar[dl]_{\alpha}&&& \pi(G) \ar[dr]^{\omega'} \ar[dl]_{\omega}\\ G\ar[rr]_{\phi} && G' &G \ar[rr]_{\phi} && G' } \]
	In particular, these two triangles imply that
	\begin{align*}
		\alpha'\circ \alpha^{-1}&=\phi \\&=\omega'\circ \omega^{-1}
	\end{align*}
\end{example}

\begin{remark}\label{rem:res} \Cref{prop:unique} can be restated as saying that, given two direct derivations $\dder{D}$ and $\dder{D'}$ with the same match, there exists an abstract equivalence between them whose first component is an identity.
\end{remark}

@@

Notice, moreover, that if $\der{D}$ is not empty, then the previous equations become
\[(\der{D}, \alpha, \omega) \equiv^a (\phi * \der{D}, \phi^{-1}\circ \alpha,   \omega) \qquad (\der{D}, \alpha, \omega) \equiv^a ( \der{D}*\psi, \alpha, \psi \circ \omega )\]
\end{remark}

@@

\begin{remark}\label{rem:lgt}
	Let $(\der{D}, \alpha, \omega)$ and $(\der{D}', \alpha', \omega')$ two composable decorated derivations   such that $\lgh(\der{D})=n$ and $\lgh(\der{D}')=m$.	 Then $(\der{D}*\omega^{-1}\cdot \alpha'*\der{D'}, \alpha, \omega')$ has length $n+m$.
\end{remark}

@@

	\begin{itemize}
		\item $\lgh(\der{D})=0$. Then, $\lgh(\der{D}')$ is $0$ too. By \Cref{def:conc} we have
		\begin{align*}
			(\der{D}, \alpha, \omega)\cdot (\der{E}, \beta, \xi)&=(\der{E}, \beta\circ \omega^{-1}\circ \alpha, \xi)\\
			(\der{D}', \alpha', \omega')\cdot (\der{E}', \beta', \xi')&=(\der{E}', \beta'\circ (\omega')^{-1}\circ \alpha', \xi')
		\end{align*}
		Now, notice that $G$ and $H$ must coincide. Moreover, by \Cref{rem:empty} we also know that $\pi(G)=\pi(G')$ too. The same \Cref{rem:empty} entails that the inner squares of the following diagram are commutative, so that the whole rectangle commutes too.
		\[\xymatrix@C=35pt{\pi(G) \ar[r]^{\alpha} \ar[d]_{\id{\pi(G)}}& G  \ar[d]_{\phi_G}\ar[r]^-{\omega^{-1}} & \pi(G) \ar[d]_{\id{\pi(G)}} \ar[r]^{\beta}& E \ar[d]^{\varphi_E}\\\pi(G') \ar[r]_{\alpha'}& G' \ar[r]_-{(\omega')^{-1}} & \pi(G') \ar[r]_{\beta'} & E'}\]

@@

		\item $\lgh(\der{D})\neq 0$ and $\lgh(\der{E})= 0$. As in the point above, we get that also $\der{E}'$  is an empty derivation, thus we have
		\begin{align*}
			(\der{D}, \alpha, \omega)\cdot (\der{E}, \beta, \xi)&=(\der{D},  \alpha, \omega \circ \beta^{-1}\circ \xi)\\
			(\der{D}', \alpha', \omega')\cdot (\der{E}', \beta', \xi')&=(\der{D}',  \alpha', \omega' \circ (\beta')^{-1}\circ \xi')
		\end{align*}
		In this case we have that $E=K$ and that $\pi(E)=\pi(E')$. From \Cref{rem:empty} we deduce that the diagram below commutes.
		\[\xymatrix@C=35pt{\pi(E) \ar[r]^{\xi} \ar[d]_{\id{\pi(E)}}& E  \ar[d]_{\varphi_E}\ar[r]^-{\beta^{-1}} & \pi(E) \ar[d]_{\id{\pi(E)}} \ar[r]^{\omega}& H \ar[d]^{\phi_H}\\\pi(E') \ar[r]_{\xi'}& E' \ar[r]_-{(\beta')^{-1}} & \pi(E') \ar[r]_{\omega'} & H'}\]
		The thesis now follows at once.
		\item $\lgh(\der{D})\neq 0$ and $\lgh(\der{E})\neq 0$. In this case we have
		\begin{align*}
			(\der{D}, \alpha, \omega)\cdot (\der{E}, \beta, \xi)&=(\der{D}*\omega^{-1}\cdot \beta*\der{E}, \alpha, \xi)\\
			(\der{D}', \alpha', \omega')\cdot (\der{E}', \beta', \xi')&=(\der{D}'*(\omega')^{-1}\cdot \beta'*\der{E'}, \alpha', \xi')
		\end{align*}
		To fix the notation, suppose that $\der{D}$, $\der{D}'$, $\der{E}$ and $\der{E}'$ are given by
		\[\der{D}=\{\dder{D}_i\}_{i=0}^n \quad \der{D}'=\{\dder{D}'_i\}_{i=0}^n \quad \der{E}=\{\dder{E}_i\}_{i=0}^t \quad \der{E}'=\{\dder{E}'_i\}_{i=0}^t\]
		Moreover, noticing that the rule applied by $\dder{D}_i$ and the one applied in $\mathcal{E}_i$  must coincide with, respectively, the one applied in $\dder{D}'_i$ and the one applied $\dder{E}'_i$. We will also assume that $\dder{D}_i$, $\dder{D}'_i$, $\dder{E}_i$ and $\dder{E}'_i$ are given, respectively, by the following four diagrams. 
		\[\xymatrix{L_{\der{D},i} \ar[d]_{m_{\der{D}, i}}& K_{\der{D},i} \ar[d]_{k_{\der{D}, i}} \ar[r]^{r_{\der{D},i}} \ar@{>->}[l]_{l_{\der{D},i}} & R_{\der{D},i}\ar[d]^{h_{\der{D}, i}} &L_{\der{E},i} \ar[d]_{m_{\der{E}, i}}& K_{\der{E},i} \ar[d]_{k_{\der{E}, i}} \ar[r]^{r_{\der{E},i}} \ar@{>->}[l]_{l_{\der{E},i}} & R_{\der{E},i}\ar[d]^{h_{\der{E}, i}} \\G_i & D_i \ar[r]_{g_{\der{D},i}} \ar@{>->}[l]^{f_{\der{D},i}} & G_{i+1} & E_i & F_i\ar[r]_{g_{\der{E},i}} \ar@{>->}[l]^{f_{\der{E},i}}  & E_{i+1}}\]
		\[\xymatrix{L_{\der{D},i} \ar[d]_{m_{\der{D}', i}}& K_{\der{D},i} \ar[d]_{k_{\der{D}', i}} \ar[r]^{r_{\der{D},i}} \ar@{>->}[l]_{l_{\der{D},i}} & R_{\der{D},i}\ar[d]^{h_{\der{D}', i}} &L_{\der{E},i} \ar[d]_{m_{\der{E}', i}}& K_{\der{E},i} \ar[d]_{k_{\der{E}', i}} \ar[r]^{r_{\der{E},i}} \ar@{>->}[l]_{l_{\der{E},i}} & R_{\der{E},i}\ar[d]^{h_{\der{E}', i}} \\G'_i & D'_i \ar[r]_{g_{\der{D}',i}} \ar@{>->}[l]^{f_{\der{D}',i}} & G'_{i+1} & E'_i & F'_i\ar[r]_{g_{\der{E}',i}} \ar@{>->}[l]^{f_{\der{E}',i}}  & E'_{i+1}}\]

@@

		The thesis now follows.	\qedhere 
	\end{itemize}	
\end{proof}

>>>>

\begin{definition}Let $(\X, \R)$ be a left-linear DPO-rewriting system,  an \emph{abstraction equivalence} between two derivations $\der{D}$ and $\der{D'}$ with the same length and such that $r(\der{D})=r(\der{D}')$, is a family of isomorphisms $\{\phi_X\}_{X\in \Deltamin(\der{D})}$ such that, for every $i\in [0, \lgh(\der{D})-1]$ the following diagram commutes
	\[\xymatrix@C=40pt{G'_i&D'_i \ar[r]^{g'_i} \ar@{>->}[l]_{f'_i}&G'_{i+1}\\  L_i \ar[u]^{m'_i} \ar[d]_{m_i}& K_i \ar[u]^{k'_i} \ar[d]_{k_i} \ar[r]^{r_i} \ar@{>->}[l]_{l_i} &R_i\ar[u]^{h'_i} \ar[d]_{h_i}\\G_i \ar@/_.45cm/[uu]_(.35){\phi_{G_i}}|\hole&D_i\ar@{>->}[l]^{f_i}\ar@/_.45cm/[uu]_(.35){\phi_{D_i}}|\hole \ar[r]_{g_i}&G_{i+1}\ar@/_.45cm/[uu]_{\phi_{H_i}}}\]

	If such an abstraction equivalence exists we will say that $\der{D}$ and $\der{D}'$ are \emph{abstraction equivalent} and we will use $\equiv^{abs}$ to denote the resulting relation-

Given two decorated derivations $(\der{D}, \alpha, \omega)$  and $(\der{D}', \alpha', \omega')$, with the same length $n$ and using the same rules, an abstraction equivalence $\{\phi_X\}_{X\in \Deltamin(\der{D})}$ is \emph{left consistent} if the left diagram below commutes, while it is \emph{right consistent} if the right one does so. A \emph{consistent} abstraction equivalence  is one which is both left and right consistent.
\[\xymatrix@C=15pt{&\pi(G_0) \ar[dr]^{\alpha'} \ar[dl]_{\alpha}&&& \pi(G_{n}) \ar[dr]^{\omega'} \ar[dl]_{\omega}\\ G_0 \ar[rr]_{\phi_{G_0}} && G'_0 &G_{n} \ar[rr]_{\phi_{G_{n}}} && G'_{n} } \]

We will say that $(\der{D}, \alpha, \omega)$  and $(\der{D}', \alpha', \omega')$ are \emph{left consistently}, respectively \emph{right consistently} or \emph{consistently}, abstraction equivalent, if there is a left consistent, respectively a right consistent or a consistent, abstraction equivalence between them. We will use $\equiv^{l}$, $\equiv^{r}$ and $\equiv^{a}$ to denote the resulting relations.
\end{definition}

\begin{proposition}\label{prop:equi}
The relations $\equiv^{abs}, \equiv^l, \equiv^r$ and $\equiv^a$ are equivalences.
\end{proposition}
\begin{proof}

	 Let $\der{D}$ be a derivation, then $\{\id{X}\}_{X\in \Deltamin(\der{D})}$ is an abstraction equivalence between $\der{D}$ and itself. Moreover, starting with a decorated derivation $(\der{D}, \alpha, \omega)$, the same $\{\id{X}\}_{X\in \Deltamin(\der{D})}$  is also consistent, so that all the four relations are reflexive.

	 For simmetry, let us suppose that $\{\phi_X\}_{X\in \Deltamin(\der{D})}$ is an abstraction equivalence between $\der{D}$ and $\der{D}'$, then $\{\phi^{-1}_X\}$ defines an abstraction equivalence  between $\der{D}'$ and $\der{D}$. Indeed, for every $i\in [0, \lgh(\der{D})-1]$ we have
	 \[\begin{split}
	 	m_i&=\id{G_i}\circ m_i\\&=\phi^{-1}_{G_i}\circ \phi_{G_i}\circ m_i\\&=\phi^{-1}_{G_i}\circ m'_i
	 \end{split} \qquad \begin{split}
 	k_i&=\id{D_i}\circ k_i\\&=\phi^{-1}_{D_i}\circ \phi_{D_i}\circ k_i\\&=\phi^{-1}_{D_i}\circ k'_i
	 \end{split}\qquad \begin{split}
 	h_i&=\id{G_{i+1}}\circ h_i\\&=\phi^{-1}_{G_{i+1}}\circ \phi_{G_{i+1}}\circ h_i\\&=\phi^{-1}_{G_{i+1}}\circ h'_i
	 \end{split} \]

	Starting with two decorated derivations $(\der{D}, \alpha, \omega)$ and $(\der{D}', \alpha', \omega')$, if $\{\phi_X\}_{X\in \Deltamin(\der{D})}$  is left, respectively right, consistent  then, putting $n$ as the length of $\der{D}$, we have:
	\[\begin{split}
		\alpha&=\id{G_0}\circ \alpha\\&=\phi^{-1}_{G_0}\circ \phi_{G_0}\circ \alpha\\&=\phi^{-1}_{G_0}\circ \alpha'
	\end{split}\qquad \begin{split}
		\omega&=\id{G_n}\circ \omega\\&=\phi^{-1}_{G_n}\circ \phi_{G_n}\circ \omega\\&=\phi^{-1}_{G_n}\circ \omega'
	\end{split}\]

	Finally, let $\{\phi_X\}_{X\in \Deltamin(\der{D})}$ and $\{\psi_X\}_{X\in \Deltamin(\der{D}')}$ be abstraction equivalence witnessing $\der{D} \equiv^{abs}\der{D}'$ and $\der{D}'\equiv^{abs}\der{D}''$. Then we can consider the family $\{\psi_X\circ \phi_X\}_{X\in \Deltamin(\der{D})}$  and check that, for every $i\in [0,\lgh(\der{D})]$ we have
	 \[\begin{split}
	\psi_{G_i}\circ \phi_{G_i}\circ m_i&= \psi_{G_i}\circ m'_i\\&=m''_i 
\end{split} \qquad \begin{split}
	\psi_{D_i}\circ \phi_{D_i}\circ k_i&= \psi_{D_i}\circ k'_i\\&=k''_i
\end{split}\qquad \begin{split}
\psi_{G_{i+1}}\circ \phi_{G_{i+1}}\circ h_i&= \psi_{G_{i+1}}\circ h'_i\\&=h''_i
\end{split} \]

	For decorations, if $\{\phi_X\}_{X\in \Deltamin(\der{D})}$ and $\{\psi_X\}_{X\in \Deltamin(\der{D}')}$ are left or right consistent we have
	 \[\begin{split}
		\psi_{G_0}\circ \phi_{G_0}\circ \alpha&= \psi_{G_0}\circ \alpha'\\&=\alpha'' 
	\end{split} \qquad \begin{split}
		\psi_{G_{\lgh(\der{D})}}\circ \phi_{G_{\lgh(\der{D})}}\circ \omega&= \psi_{G_{\lgh(\der{D})}}\circ \omega'\\&=\omega''
	\end{split}\]
	And we can conclude.
\end{proof}

\begin{notation}
	Since we are mainly interested into consistent abstraction equivalence, we will use  $[\der{D}, \alpha, \omega]_a$ to denote the equivalence class of  $(\der{D}, \alpha, \omega)$ with respect to $\equiv^a$.
	Such equivalence classes will be called  \emph{abstract decorated derivations}.  
\end{notation}

\begin{example}\label{rem:empty}
	Let $\{G\}$ and  $\{G'\}'$ be two empty derivations  An abstraction equivalence between $\{G\}$ and $\{G\}'$ is just an isomorphism $\phi\colon G\to G'$, so that $\pi(G)=\pi(G')$. If we are dealing with decorated derivations $(\{G\}, \alpha, \omega)$ and $(\{G'\}, \alpha, \omega)$,  left and right consistency of $\phi $ correspond to the commutativity of the diagrams below.
	\[\xymatrix@C=15pt{&\pi(G) \ar[dr]^{\alpha'} \ar[dl]_{\alpha}&&& \pi(G) \ar[dr]^{\omega'} \ar[dl]_{\omega}\\ G\ar[rr]_{\phi} && G' &G \ar[rr]_{\phi} && G' } \]

	In particular, this prove that two empty decorated derivations $(\{G\},\alpha, \omega) $ and $(\{G'\}',\alpha', \omega')$ are consistently abstraction equivalent if and only if
	\[	\alpha'\circ \alpha^{-1}=\omega'\circ \omega^{-1}\]
\end{example}

\begin{remark}\label{rem:res} \Cref{prop:unique} can be restated as saying that, given two direct derivations $\dder{D}$ and $\dder{D'}$ with the same match, there exists an abstract equivalence between them whose first component is the identity. 
\end{remark}

@@

if $\der{D}$ is not empty, then the previous equations become
\[(\der{D}, \alpha, \omega) \equiv^a (\phi * \der{D}, \phi^{-1}\circ \alpha,   \omega) \qquad (\der{D}, \alpha, \omega) \equiv^a ( \der{D}*\psi, \alpha, \psi \circ \omega )\]

In particular, given an empty decorated derivation $(\{G\}, \alpha, \omega)$, we can apply what we have noticed to $\alpha \colon \pi(G)\to G$ and to $\omega^{-1}\colon G\to \pi(G) $ to get that 
\[(\{G\}, \alpha, \omega)\equiv^a (\{\pi(G)\}, \id{\pi(G)}, \alpha^{-1}\circ \omega) \qquad (\{G\}, \alpha, \omega)\equiv^a (\{\pi(G)\}, 	\omega^{-1}\circ \alpha, \id{\pi(G)})\]
\end{remark}

@@

\begin{remark}\label{rem:lgt}
	Let $(\der{D}, \alpha, \omega)$ and $(\der{D}', \alpha', \omega')$ two composable decorated derivations   such that $\lgh(\der{D})=n$ and $\lgh(\der{D}')=m$.	 Then $(\der{D}, \alpha, \omega)\cdot (\der{D}', \alpha', \omega')$ has length $n+m$.
\end{remark}

\begin{remark}\label{rem:id}
	Let us consider  $(\{G\}, \alpha, \alpha)$ be a decoration of length $0$ with the same isomorphism as initial and final decoration. Then for every other derivations $(\der{D}, \beta, \omega)$ with source isomorphic to $G$, we have
	\[(\{G\}, \alpha, \alpha)\cdot (\der{D}, \beta, \omega) = (\der{D}, \beta, \omega)\]

	On the other hand, if  $(\der{E}, \gamma, \zeta)$  has a target isomorphic to $G$, we have
	\[(\der{E}, \gamma, \zeta)\cdot (\{G\}, \alpha, \alpha) = \begin{cases}
		(\{G\}, \alpha \circ \zeta^{-1}\circ \gamma, \alpha)	&\lgh(\der{E})=0 \\
		(\der{E}, \gamma, \zeta )& \lgh(\der{E})= 0
	\end{cases}\]

Now,  in the first case, we have that
\[\alpha \circ \zeta^{-1} \circ \gamma \circ \gamma ^{-1} = \alpha \circ \zeta^{-1}\]
therefore, by \Cref{rem:empty}, $(\der{E}, \gamma, \zeta)\cdot (\{G\}, \alpha, \alpha)\equiv^a  (\der{E}, \gamma, \zeta)$.
\end{remark}

\begin{remark}\label{ex:zero}
Let $(\{G\}, \alpha, \omega)$ and $(\{G'\}, \alpha', \omega')$ be two decorated derivations of length zero with $G$ and $G'$ isomorphic. Then we can always compose with another empty decorated derivations to get one from the other. In particular, using \Cref{rem:absequi} we have:
\begin{align*}
(\{G\}, \alpha, \omega)&\equiv^{a} (\pi(G), \id{\pi(G)}, \alpha^{-1}\circ \omega)\\&=(\{\pi(G)\}, (\alpha')^{-1}\circ \omega'\circ \omega'^{-1} \circ \alpha', \alpha^{-1}\circ \omega)\\&=(\{G'\}, \alpha', \omega')\cdot (\{\pi(G)\}, (\alpha')^{-1}\circ \omega', \alpha^{-1}\circ \omega)
\\(\{G\}, \alpha, \omega)&\equiv^{a} (\{\pi(G)\}, 	\omega^{-1}\circ \alpha, \id{\pi(G)})\\&=(\{\pi(G)\}, \omega^{-1}\circ \alpha, (\omega')^{-1}\circ\alpha'\circ (\alpha')^{-1} \circ \omega')\\&= (\{\pi(G)\}, \omega^{-1}\circ \alpha, (\omega')^{-1}\circ\alpha') \cdot (\{G'\}, \alpha', \omega')
\end{align*}
\end{remark}

We can generalize the previous remark to the following elementary, but useful, result about abstraction equivalence of composite decorated derivations. 
\begin{lemma}\label{rem:obv}
	Let  $(\der{D}_1, \alpha_1, \omega_1)$, $(\der{D}_2, \alpha_2, \omega_2)$, $(\der{D}'_1, \alpha'_1, \omega'_1)$ and $(\der{D}'_2, \alpha'_2, \omega'_2)$ be two pairs of composable derivations and define $(\der{E}, \xi, \theta)$ and $(\der{E}', \xi', \theta')$ as 
	\[(\der{E}, \xi, \theta):=(\der{D}_1, \alpha_1, \omega_1) \cdot (\der{D}_2, \alpha_2, \omega_2) \qquad (\der{E}', \xi', \theta'):=(\der{D}'_1, \alpha'_1, \omega'_1) \cdot (\der{D}'_2, \alpha'_2, \omega'_2)\]
	 If $\lgh(\der{D}_1)=\lgh(\der{D}'_1)$, $\lgh(\der{D}_2)=\lgh(\der{D}'_2)$ and $\der{E}\equiv^a \der{E}'$  then there exist empty derivations $(\{A\}, \gamma_1, \delta_1)$ and $(\{B\}, \gamma_2, \delta_2)$ such that
	 \[(\der{D}_1, \alpha_1, \omega_1) \equiv^a (\der{D}'_1, \alpha'_1, \omega'_1) \cdot  (\{A\}, \gamma_1, \delta_1) \qquad (\der{D}_2, \alpha_2, \omega_2) \equiv^a (\{B\}, \gamma_2, \delta_2)\cdot (\der{D}'_2, \alpha'_2, \omega'_2)  \]
\end{lemma}
\begin{proof} Let $\{\phi_X\}_{X\in \Deltamin(\der{E})}$ be an abstraction equivalent witnessing that $(\der{E}, \zeta, \theta)\equiv^{a}(\der{E}', \zeta', \theta')$.	We split the cases following \Cref{def:conc}.

	\smallskip \noindent $\lgh(\der{D}_1)=0$. In this case we have that
	\[(\der{E}, \zeta, \theta)=(\der{D}_2, \alpha_2\circ \omega^{-1}_1 \circ \alpha_1, \omega_2) \qquad (\der{E}', \zeta', \theta')=(\der{D}'_2, \alpha'_2\circ (\omega'_1)^{-1} \circ \alpha'_1, \omega'_2)\]
	By hypothesis $\der{D}_1$ and $\der{D}'_1$ are given by $\{G_{1,0}\}$ and $\{G'_{1,0}\}$. Moreover we know that $G_{1,0}$ is isomorphic to $G_{2,0}$ and  $G'_{1,0}$ to $G'_{2,0}$. Since $\phi_{G_{2,0}}$ is an isomorphism we conclude that $G_{1,0}$ and $G_{2,0}$ are isomorphic and we deduce the first half of the thesis from \Cref{ex:zero}.

	For the second half, let us consider define $B$ as $\pi(G_{2,0})$ and
	\[\gamma_2:=\alpha^{-1}_1\circ \omega_1\qquad  \delta_2:=(\alpha'_1)^{-1}\circ \omega'_1  \]

	Hence we have
\[
(\{\pi(G_{2,0})\}, \alpha^{-1}_1\circ \omega_1,  (\alpha'_1)^{-1}\circ \omega'_1 )\cdot(\der{D}'_2, \alpha'_2, \omega'_2)  =(\der{D}_2, \alpha'_2\circ (\omega'_1)^{-1} \circ \alpha'_1\circ \alpha_1^{-1}\circ \omega_1, \omega'_2)\]
Now, by consistency of $\{\phi_{X}\}_{X\in \Deltamin(\der{D}_2)}$ we know that
\[\phi_{G_{2,0}}\circ \alpha_2\circ \omega^{-1}_1\circ \alpha_1=\alpha'_2\circ (\omega'_1)^{-1} \circ \alpha'_1\]
	therefore we also have the equality
	\[\phi_{G_{2,0}}\circ \alpha_2=\alpha'_2\circ (\omega'_1)^{-1} \circ \alpha'_1 \circ \alpha^{-1}_1\circ \omega_1 \]
	We conclude that $\{\phi_X\}_{X\in \der{D}_2}$ wintesses the thesis.

	\smallskip \noindent $\lgh(\der{D}_2)=0$ and $\lgh(\der{D}_1)\neq 0$. By definition we know that
	\[(\der{E}, \zeta, \theta)=(\der{D}_1, \alpha_1, \omega_1\circ \alpha_2^{-1}\circ \omega_2) \qquad (\der{E}', \zeta', \theta')=(\der{D}'_1, \alpha'_1, \omega'_1\circ (\alpha'_2)^{-1}\circ \omega'_2)\]

	Let $l$ be the length of $\der{D}_1$, and consider $(\{\pi(G'_{1, l})\}, (\omega'_2)^{-1}\circ \alpha'_2, \omega^{-1}_2\circ \alpha_2)$, so that
	\[ (\der{D}'_1, \alpha'_1, \omega'_1) \cdot  (\{\pi(G_{1, l})\}, (\omega'_2)^{-1}\circ \alpha'_2, \omega^{-1}_2\circ \alpha_2)= (\der{D}'_1, \alpha'_1, \omega'_1\circ (\alpha'_2)^{-1}\circ \omega'_2 \circ \omega^{-1}_2\circ \alpha_2)\]
	Now, let us consider the abstraction equivalence between  the decorated derivations $(\der{D}_1, \alpha_1, \omega_1)$ and  $(\der{D}'_1, \alpha'_1, \omega'_1) \cdot  (\{\pi(G_{1, l})\}, (\omega'_2)^{-1}\circ \alpha'_2, \omega^{-1}_2\circ \alpha_2)$ given by $\{\phi_X\}_{X\in \Deltamin(\der{D}_1)}$. On the one hand it is, by construction, left consistent.  On the other hand we know that

	\[\phi_{G_{1,l}}\circ \omega_1\circ \alpha_2^{-1}\circ \omega_2=\omega'_1\circ (\alpha'_2)^{-1}\circ \omega'_2 \]

	 Composing on both sides with $\omega^{-1}_1\circ \alpha_2$ yields now the thesis.

The second half of the thesis follows immediately from \Cref{ex:zero}.

	\smallskip \noindent  $\lgh(\der{D}_1)$ and $\lgh(\der{D}_2)$ are both different from $0$. By \Cref{def:conc} in this case we have
	\[(\der{E}, \zeta, \theta)=(\der{D}_1*\omega^{-1}_1 \cdot \alpha_2*\der{D}_2, \alpha_1,\omega_2 ) \qquad (\der{E}', \zeta', \theta')=(\der{D}'_1*(\omega'_1)^{-1} \cdot \alpha'_2*\der{D}'_2, \alpha'_1,\omega'_2 )\]

\end{proof}

@@

	\smallskip \noindent  $\lgh(\der{D})=0$. Then, $\lgh(\der{D}')$ is $0$ too. By \Cref{def:conc} we have
		\begin{align*}
			(\der{D}, \alpha, \omega)\cdot (\der{E}, \beta, \xi)&=(\der{E}, \beta\circ \omega^{-1}\circ \alpha, \xi)\\
			(\der{D}', \alpha', \omega')\cdot (\der{E}', \beta', \xi')&=(\der{E}', \beta'\circ (\omega')^{-1}\circ \alpha', \xi')
		\end{align*}
		Now, notice that $G$ and $H$ must coincide. Moreover, by \Cref{rem:empty} we also know that $\pi(G)=\pi(G')$ too. The same \Cref{rem:empty} entails that the inner squares of the following diagram are commutative, so that the whole rectangle commutes too.
		\[\xymatrix@C=35pt{\pi(G) \ar[r]^{\alpha} \ar[d]_{\id{\pi(G)}}& G  \ar[d]_{\phi_G}\ar[r]^-{\omega^{-1}} & \pi(G) \ar[d]_{\id{\pi(G)}} \ar[r]^{\beta}& E \ar[d]^{\varphi_E}\\\pi(G') \ar[r]_{\alpha'}& G' \ar[r]_-{(\omega')^{-1}} & \pi(G') \ar[r]_{\beta'} & E'}\]

@@

\smallskip \noindent  $\lgh(\der{D})\neq 0$ and $\lgh(\der{E})= 0$. As in the point above, we get that also $\der{E}'$  is an empty derivation, thus we have
		\begin{align*}
			(\der{D}, \alpha, \omega)\cdot (\der{E}, \beta, \xi)&=(\der{D},  \alpha, \omega \circ \beta^{-1}\circ \xi)\\
			(\der{D}', \alpha', \omega')\cdot (\der{E}', \beta', \xi')&=(\der{D}',  \alpha', \omega' \circ (\beta')^{-1}\circ \xi')
		\end{align*}
		In this case we have that $E=K$ and that $\pi(E)=\pi(E')$. From \Cref{rem:empty} we deduce that the diagram below commutes.
		\[\xymatrix@C=35pt{\pi(E) \ar[r]^{\xi} \ar[d]_{\id{\pi(E)}}& E  \ar[d]_{\varphi_E}\ar[r]^-{\beta^{-1}} & \pi(E) \ar[d]_{\id{\pi(E)}} \ar[r]^{\omega}& H \ar[d]^{\phi_H}\\\pi(E') \ar[r]_{\xi'}& E' \ar[r]_-{(\beta')^{-1}} & \pi(E') \ar[r]_{\omega'} & H'}\]
		The thesis now follows at once.

		\smallskip \noindent $\lgh(\der{D})\neq 0$ and $\lgh(\der{E})\neq 0$. In this case we have
		\begin{align*}
			(\der{D}, \alpha, \omega)\cdot (\der{E}, \beta, \xi)&=(\der{D}*\omega^{-1}\cdot \beta*\der{E}, \alpha, \xi)\\
			(\der{D}', \alpha', \omega')\cdot (\der{E}', \beta', \xi')&=(\der{D}'*(\omega')^{-1}\cdot \beta'*\der{E'}, \alpha', \xi')
		\end{align*}
		To fix the notation, suppose that $\der{D}$, $\der{D}'$, $\der{E}$ and $\der{E}'$ are given by
		\[\der{D}=\{\dder{D}_i\}_{i=0}^n \quad \der{D}'=\{\dder{D}'_i\}_{i=0}^n \quad \der{E}=\{\dder{E}_i\}_{i=0}^t \quad \der{E}'=\{\dder{E}'_i\}_{i=0}^t\]
		Moreover, noticing that the rule applied by $\dder{D}_i$ and the one applied in $\mathcal{E}_i$  must coincide with, respectively, the one applied in $\dder{D}'_i$ and the one applied $\dder{E}'_i$. We will also assume that $\dder{D}_i$, $\dder{D}'_i$, $\dder{E}_i$ and $\dder{E}'_i$ are given, respectively, by the following four diagrams. 
		\[\xymatrix{L_{\der{D},i} \ar[d]_{m_{\der{D}, i}}& K_{\der{D},i} \ar[d]_{k_{\der{D}, i}} \ar[r]^{r_{\der{D},i}} \ar@{>->}[l]_{l_{\der{D},i}} & R_{\der{D},i}\ar[d]^{h_{\der{D}, i}} &L_{\der{E},i} \ar[d]_{m_{\der{E}, i}}& K_{\der{E},i} \ar[d]_{k_{\der{E}, i}} \ar[r]^{r_{\der{E},i}} \ar@{>->}[l]_{l_{\der{E},i}} & R_{\der{E},i}\ar[d]^{h_{\der{E}, i}} \\G_i & D_i \ar[r]_{g_{\der{D},i}} \ar@{>->}[l]^{f_{\der{D},i}} & G_{i+1} & E_i & F_i\ar[r]_{g_{\der{E},i}} \ar@{>->}[l]^{f_{\der{E},i}}  & E_{i+1}}\]
		\[\xymatrix{L_{\der{D},i} \ar[d]_{m_{\der{D}', i}}& K_{\der{D},i} \ar[d]_{k_{\der{D}', i}} \ar[r]^{r_{\der{D},i}} \ar@{>->}[l]_{l_{\der{D},i}} & R_{\der{D},i}\ar[d]^{h_{\der{D}', i}} &L_{\der{E},i} \ar[d]_{m_{\der{E}', i}}& K_{\der{E},i} \ar[d]_{k_{\der{E}', i}} \ar[r]^{r_{\der{E},i}} \ar@{>->}[l]_{l_{\der{E},i}} & R_{\der{E},i}\ar[d]^{h_{\der{E}', i}} \\G'_i & D'_i \ar[r]_{g_{\der{D}',i}} \ar@{>->}[l]^{f_{\der{D}',i}} & G'_{i+1} & E'_i & F'_i\ar[r]_{g_{\der{E}',i}} \ar@{>->}[l]^{f_{\der{E}',i}}  & E'_{i+1}}\]

@@

		The thesis now follows.	
\end{proof}

We are now ready to prove that the operatione defined by \Cref{def:conc} is associative up to decorated abstraction equivalence.
